CF Registry Analysis

### 2022-09-07 Christine Chan/CF Registry/cf_registry_analysis.Rmd
Update cf_registry_analysis.Rmd
tidy formatting

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Christine Chan/CF Registry")
```

```{r}
# Read in CSVs
df = read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv")
annual = read.csv("./Data_Raw/CFF19_Annualized_Del1.csv")
colnames(annual)[colnames(annual)=="ReviewYear"] = "reviewyear"
demo = read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
# Combine and format demographics columns 
demo$Race = apply(demo,1,function(r){
  race = r[paste0("Race",1:6)]
  race = as.numeric(which(race == 1))
  race = paste(race,collapse = ", ")
})
demo$Race[grep(", ",demo$Race)] = "More Than One"
demo$Race = factor(demo$Race,levels = c(1:6,"More Than One"),
                   labels = c("White","Black or African American",
                              "American Indian or Alaska Native","Asian",
                              "Native Hawaiian or Other Pacific Islander",
                              "Some other race","More Than One"))
demo$Hispanicrace = factor(demo$Hispanicrace,levels = 1:2,
                           labels = c("Yes","No"))
# Merge
df = left_join(df,annual, by=c('eDWID', 'reviewyear'))
df = left_join(df,demo,by = "eDWID")
# Group years by 2003-2009, 2010-2013, and 2014-2018
df$Period = cut(df$reviewyear,
                breaks =c(2002, 2009, 2013, 2018),
                labels = c("2003 - 2009","2010 - 2013","2014 - 2018"))
df$smoking = factor(df$smoking,levels = 1:4,labels = c("No",
                                                       "Occasionally",
                                                       "Yes, Regularly, less than 1 ppd",
                                                       "Yes, Regularly, 1 ppd or more"),ordered = T)
df$second_smoke = factor(df$smoking,levels = 1:4,labels = c("No",
                                                            "Occasionally",
                                                            "Yes, Regularly, less than 1 ppd",
                                                            "Yes, Regularly, 1 ppd or more"),ordered = T)
# Factors
ed_levels = c("Less than High School",
              "High School diploma or equivalent",
              "Some College",
              "College Graduate",
              "Masters/Doctoral level degree")
df$patient_education = factor(df$patient_education,levels = c(1:5),
                              labels = ed_levels)
df$father_education = factor(df$father_education,levels = c(1:5),
                             labels = ed_levels)
df$mother_education = factor(df$mother_education,levels = c(1:5),
                             labels = ed_levels)
df$spouse_education = factor(df$spouse_education,levels = c(1:5),
                             labels = ed_levels)
df$pregnant = factor(df$pregnant,levels = 0:1,labels = c("No","Yes"),ordered = T)
# Combine education checkboxes
df = df[1:1000,]
df$employment = apply(df,1,function(r){
  emp = r[paste0("employment",1:7)]
  emp = as.numeric(emp)
  if(all(is.na(emp))){
    return(NA)
  } else {
    emp = as.numeric(emp)
    emp = which(emp == 1)
    emp = paste(emp,collapse = ", ")
    return(emp)
  }
})
df$employment = sub(sub("1","Part Time",df$employment))
df$employment = sub(sub("2","Full time homemaker",df$employment))
df$employment = sub(sub("3","Full time employment",df$employment))
df$employment = sub(sub("4","Unemployed",df$employment))
df$employment = sub(sub("5","Student",df$employment))
df$employment = sub(sub("6","Disabled",df$employment))
df$employment = sub(sub("7","Retired",df$employment))
# Small dataframe for checking
test = df[1:1000,]
write.csv(test,"~/test.csv",row.names = F)
```

@@

```{r warning=FALSE}
t1_df = df %>% group_by(eDWID,Period) %>% 
  summarise(Age = max(Age_YrEnd,na.rm = T),Race = last(na.omit(Race)),
            Hispanic = last(na.omit(Hispanicrace)),Gender = last(na.omit(Gender)),
            `Patient Education` = last(na.omit(patient_education)),
            `Father Education` = last(na.omit(father_education)),
            `Mother Education` = last(na.omit(mother_education)),
            `Spouse Education` = last(na.omit(spouse_education)),
            `Marital Status` = last(na.omit(marital_status)),
            Employment = last(na.omit(employment)),
            Pregnancy = max(pregnant,na.rm = T)
  )
# Replace infinite values with missing
t1_df = data.frame(lapply(t1_df, function(c){
  c[is.infinite(c)] = NA
  return(c)
}))

cols = colnames(t1_df)[3:ncol(t1_df)]
t1 = tableby(as.formula(paste0("Period~",paste0("`",cols,"`",collapse = "+"))),
             data = t1_df)
```

>>>>

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(parallel)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Christine Chan/CF Registry")
```

```{r}
# Read in CSVs
df <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv")
annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv")
colnames(annual)[colnames(annual) == "ReviewYear"] <- "reviewyear"
demo <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
# Combine and format demographics columns
demo$Race <- apply(demo, 1, function(r) {
  race <- r[paste0("Race", 1:6)]
  race <- as.numeric(which(race == 1))
  race <- paste(race, collapse = ", ")
})
demo$Race[grep(", ", demo$Race)] <- "More Than One"
demo$Race <- factor(demo$Race,
  levels = c(1:6, "More Than One"),
  labels = c(
    "White", "Black or African American",
    "American Indian or Alaska Native", "Asian",
    "Native Hawaiian or Other Pacific Islander",
    "Some other race", "More Than One"
  )
)
demo$Hispanicrace <- factor(demo$Hispanicrace,
  levels = 1:2,
  labels = c("Yes", "No")
)
# Merge
df <- left_join(df, annual, by = c("eDWID", "reviewyear"))
df <- left_join(df, demo, by = "eDWID")
# Group years by 2003-2009, 2010-2013, and 2014-2018
df$Period <- cut(df$reviewyear,
  breaks = c(2002, 2009, 2013, 2018),
  labels = c("2003 - 2009", "2010 - 2013", "2014 - 2018")
)
df$smoking <- factor(df$smoking, levels = 1:4, labels = c(
  "No",
  "Occasionally",
  "Yes, Regularly, less than 1 ppd",
  "Yes, Regularly, 1 ppd or more"
), ordered = T)
df$second_smoke <- factor(df$smoking, levels = 1:4, labels = c(
  "No",
  "Occasionally",
  "Yes, Regularly, less than 1 ppd",
  "Yes, Regularly, 1 ppd or more"
), ordered = T)
# Factors
ed_levels <- c(
  "Less than High School",
  "High School diploma or equivalent",
  "Some College",
  "College Graduate",
  "Masters/Doctoral level degree"
)
df$patient_education <- factor(df$patient_education,
  levels = c(1:5),
  labels = ed_levels
)
df$father_education <- factor(df$father_education,
  levels = c(1:5),
  labels = ed_levels
)
df$mother_education <- factor(df$mother_education,
  levels = c(1:5),
  labels = ed_levels
)
df$spouse_education <- factor(df$spouse_education,
  levels = c(1:5),
  labels = ed_levels
)
df$pregnant <- factor(df$pregnant, levels = 0:1, labels = c("No", "Yes"), ordered = T)
# Combine education checkboxes
cl <- makeCluster(detectCores() * 0.5, type = "FORK")
df$employment <- parApply(cl, df, 1, function(r) {
  emp <- r[paste0("employment", 1:7)]
  emp <- as.numeric(emp)
  if (all(is.na(emp) | emp == 0)) {
    return(NA)
  } else {
    emp <- as.numeric(emp)
    emp <- which(emp == 1)
    emp <- paste(emp, collapse = ", ")
    return(emp)
  }
})
stopCluster(cl)
df$employment <- sub("1", "Part Time", df$employment)
df$employment <- sub("2", "Full time homemaker", df$employment)
df$employment <- sub("3", "Full time employment", df$employment)
df$employment <- sub("4", "Unemployed", df$employment)
df$employment <- sub("5", "Student", df$employment)
df$employment <- sub("6", "Disabled", df$employment)
df$employment <- sub("7", "Retired", df$employment)
df$employment <- as.factor(df$employment)
# Small dataframe for checking
test <- df[1:1000, ]
write.csv(test, "~/test.csv", row.names = F, na = "")
```

@@

```{r warning=FALSE,results='asis'}
# Summarize by ID and period
t1_df <- df %>%
  group_by(eDWID, Period) %>%
  summarise(
    Age = max(Age_YrEnd), Race = last(na.omit(Race)),
    Hispanic = last(na.omit(Hispanicrace)), Gender = last(na.omit(Gender)),
    `Patient Education` = last(na.omit(patient_education)),
    `Father Education` = last(na.omit(father_education)),
    `Mother Education` = last(na.omit(mother_education)),
    `Spouse Education` = last(na.omit(spouse_education)),
    `Marital Status` = last(na.omit(marital_status)),
    Employment = last(na.omit(employment)),
    Pregnancy = max(pregnant, na.rm = T)
  )
# Replace infinite values with missing
t1_df <- data.frame(lapply(t1_df, function(c) {
  c[is.infinite(c)] <- NA
  return(c)
}))
# Print table 1
cols <- colnames(t1_df)[3:ncol(t1_df)]
t1 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
  data = t1_df
)
summary(t1)
```
